---
output:
  md_document:
    variant: gfm
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%",
    message = FALSE,
    warning = FALSE
)
```

```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
```

# ggsector: Easily draw sectors with grid and ggplot2

<!-- badges: start -->
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
<!-- [![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/ggsector)](https://cran.r-project.org/package=ggsector) -->
`r badge_devel("yanpd01/ggsector", "green")`

`r badger::badge_custom("platforms", "all", "green")`
`r badger::badge_custom("Windows", "passing", "green")`
`r badger::badge_custom("Linux", "passing", "green")`
<!-- badges: end -->

```{r comment="", echo=FALSE, results='asis'}
cat(packageDescription("ggsector")$Description)
```

## :writing_hand: Authors

Pengdong Yan



## :arrow_double_down: Installation

Get the released version from Cran:

```{r, eval = FALSE}
## The new version of the R package is still under review.
## If you want to try it out, you can use the following github method to install it
# install.packages("ggsector")
```

Or the development version from github or gitee:

```{r, eval = FALSE}
## install.packages("remotes")
# from github
remotes::install_github(
    "yanpd01/ggsector",
    build_vignettes = TRUE
)
# from gitee
remotes::install_git(
    "https://gitee.com/yanpd01/ggsector",
    build_vignettes = TRUE
)
```


##  :books: Usage

For the usage of this R package, please type `vignette("ggsector")` after installation to view it.
There are detailed case descriptions in it.


## :sparkling_heart: Acknowledgments

The code of this R package refers to [jjplot](https://github.com/junjunlab/jjPlot) of [JunJunLao](https://github.com/junjunlab) and [ggplot2](https://github.com/tidyverse/ggplot2) of [Hadley](https://github.com/hadley).

The Description, vignette, and readme of this R package refer to [clusterProfiler](https://github.com/YuLab-SMU/clusterProfiler), [ggfun](https://github.com/YuLab-SMU/ggfun), and [treeio](https://github.com/YuLab-SMU/treeio) of [Guangchuang YU](https://github.com/YuLab-SMU/).

Here, I would like to express my highest respect to thank the big guys for their open source spirit.


