---
title: "ggsector"
author: "Pengdong Yan"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
  pdf_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{ggsector}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 8,
    fig.height = 8
)
```

Load the required R packages.

```{r setup, message = FALSE}
library(magrittr)
library(grid)
library(ComplexHeatmap)
library(Seurat)
library(ggplot2)
library(ggsector)
```


## Use original coordinates with `grid.polygon`

### coordinates of single sector

type of percent, start = 0, r_start = 0

```{r, fig.width=3, fig.height=3}
tmp_df <- sector_df(x = 0.5, y = 0.5, theta = 25, r = 0.4, start = 0, r_start = 0)
tmp_df
grid.newpage()
grid.polygon(
    tmp_df$x, tmp_df$y,
    vp = viewport(height = unit(1, "snpc"), width = unit(1, "snpc"))
)
```

type of percent, start = 50, r_start = 0.2

```{r, fig.width=3, fig.height=3}
tmp_df <- sector_df(x = 0.5, y = 0.5, theta = 25, r = 0.4, start = 50, r_start = 0.2)
tmp_df
grid.newpage()
grid.polygon(
    tmp_df$x, tmp_df$y,
    vp = viewport(height = unit(1, "snpc"), width = unit(1, "snpc"))
)
```

type of degree, start = 90, r_start = 0

```{r, fig.width=3, fig.height=3}
tmp_df <- sector_df(
    x = 0.5, y = 0.5, theta = 180, r = 0.4,
    start = 90, r_start = 0, type = "degree"
)
tmp_df
grid.newpage()
grid.polygon(
    tmp_df$x, tmp_df$y,
    vp = viewport(height = unit(1, "snpc"), width = unit(1, "snpc"))
)
```

type of degree, start = 180, r_start = 0.2

```{r, fig.width=3, fig.height=3}
tmp_df <- sector_df(
    x = 0.5, y = 0.5, theta = 180, r = 0.4,
    start = 270, r_start = 0.2, type = "degree"
)
tmp_df
grid.newpage()
grid.polygon(
    tmp_df$x, tmp_df$y,
    vp = viewport(height = unit(1, "snpc"), width = unit(1, "snpc"))
)
```

### Coordinates of Multiple Sectors


```{r, fig.width=3, fig.height=3}
tmp_df <- sector_df_multiple(
    x = c(0.2, 0.5, 0.8),
    theta = c(25, 50, 75),
    r = 0.15,
    start = c(75, 50, 100),
    r_start = c(0, 0.05, 0.1),
    type = "percent"
)
tmp_df
grid.newpage()
grid.polygon(
    tmp_df$x,
    tmp_df$y,
    id = tmp_df$group,
    vp = viewport(height = unit(1, "snpc"), width = unit(1, "snpc")),
    gp = gpar(
        fill = 3:1, col = 1:3
    )
)
```


## Use ggsector with grid

`sectorGrob` with units of "cm" and type of "degree"

```{r, fig.width=5, fig.height=4}
grid.newpage()
gp <- sectorGrob(
    x = unit(c(3, 9, 15), "cm"),
    y = unit(c(5, 9, 15), "cm"),
    theta = c(90, 180, 270),
    r = 1,
    start = c(180, 180, 270),
    r_start = c(0.6, 0.3, 0),
    type = "degree",
    group = factor(1:3, levels = c(2, 3, 1)),
    gp = gpar(fill = c("green", "red", "blue"))
)
grid.draw(gp)
```

`grid.sector` with units of "npc" and type of "percent"

```{r, fig.width=5, fig.height=4}
grid.newpage()
grid.sector(
    x = c(0.1, 0.5, 0.9),
    y = c(0.9, 0.6, 0.1),
    theta = c(25, 50, 90),
    r = .1,
    start = c(25, 50, 100),
    r_start = c(0.06, 0.03, 0),
    type = "percent",
    group = factor(1:3, levels = c(2, 3, 1)),
    gp = gpar(col = c("green", "red", "blue"), fill = 2:4),
    default.units = "npc"
)
```

## Use ggsector with ComplexHeatmap

Prepare data

```{r}
library(magrittr)
library(ComplexHeatmap)
t0 <- cor(mtcars) %>%
    set_colnames(paste("y_", colnames(.))) %>%
    set_rownames(paste("x_", rownames(.)))
mat <- abs(t0)
mat[1:5, 1:5]
```

Realized by modifying the [grid::viewport()] with cell_fun, the sector can be set with a fixed width and height

```{r}
set.seed(1)
Heatmap(
    mat,
    name = "vp",
    rect_gp = gpar(type = "none"),
    cell_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(
            x = x, y = y, width = width, height = height,
            gp = gpar(col = "grey", fill = NA)
        )
        grid.sector(
            theta = mat[i, j] * 100,
            r = 0.5,
            start = mat[i, j] * 100 * runif(1),
            r_start = mat[i, j] * 0.49 * runif(1),
            vp = viewport(x, y, width, height),
            gp = gpar(fill = fill, col = "transparent")
        )
    },
    width = unit(.7, "snpc"),
    height = unit(.7, "snpc")
)
```

Realized in the form of coordinates + radius with cell_fun.

```{r}
# The default viewport locks the horizontal and vertical axes
# so that the sector does not deform, which needs to be removed here.
# The radius 'r' is half the min(length, width).
set.seed(2)
Heatmap(
    mat,
    name = "xy + r",
    rect_gp = gpar(type = "none"),
    cell_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(
            x = x, y = y, width = width, height = height,
            gp = gpar(col = "grey", fill = NA)
        )
        r <- as.numeric(min(width, height)) / 2
        grid.sector(
            x,
            y,
            theta = mat[i, j] * 100,
            r = r,
            start = mat[i, j] * 100 * runif(1),
            r_start = mat[i, j] * r * 0.9 * runif(1),
            vp = NULL,
            gp = gpar(fill = fill, col = "transparent")
        )
    },
    width = unit(.7, "snpc"),
    height = unit(.7, "snpc")
)
```

Realized With layer full

```{r}
# The input matrix needs to be extracted with pindex(mat, i, j)
set.seed(3)
Heatmap(
    mat,
    name = "layer",
    rect_gp = gpar(type = "none"),
    layer_fun = function(j, i, x, y, width, height, fill) {
        grid.rect(
            x = x, y = y, width = width, height = height,
            gp = gpar(col = "grey", fill = NA)
        )
        r <- as.numeric(min(width, height)) / 2
        grid.sector(
            x,
            y,
            theta = pindex(mat, i, j) * 100,
            r = r,
            start = pindex(mat, i, j) * 100 * runif(nrow(mat) * ncol(mat)),
            r_start = pindex(mat, i, j) * r * 0.9 * runif(nrow(mat) * ncol(mat)),
            vp = NULL,
            gp = gpar(fill = fill, col = "transparent")
        )
    },
    width = unit(.7, "snpc"),
    height = unit(.7, "snpc")
)
```

## Use ggsector with ggplot2

Prepare data

```{r}
set.seed(1)
t0 <- cor(mtcars) %>%
    set_colnames(paste("y_", colnames(.))) %>%
    set_rownames(paste("x_", rownames(.)))
t1 <- t0 %>%
    reshape2::melt()
t1$group <- sample(letters[1:5], nrow(t1), TRUE)
t1$r <- sample(c(0.24, 0.48), nrow(t1), TRUE)
t1$start <- runif(nrow(t1), max = 50)
t1$r_start <- sample(c(0, 0.08, 0.16), nrow(t1), TRUE)
t1$v <- abs(t1$value)
head(t1)
```

Plot with type of "percent"

```{r}
ggplot(t1) +
    aes(x = Var1, y = Var2, theta = v * 100) +
    geom_sector(
        aes(fill = group, r = r, start = start, r_start = r_start),
        color = "transparent"
    ) +
    theme_bw()
```

Plot with type of "degree" and coord_fixed()

```{r}
ggplot(t1) +
    aes(x = Var1, y = Var2, theta = v * 360) +
    geom_sector(
        aes(fill = group, r = r, start = start * 3.6, r_start = r_start),
        color = "transparent",
        type = "degree"
    ) +
    coord_fixed() +
    theme_bw()
```

Draw individually

```{r}
ggplot(t1) +
    aes(x = Var1, y = Var2, theta = v * 100, r_start = r_start) +
    geom_sector(
        aes(fill = group, r = r, start = start),
        color = "transparent",
        individual = TRUE
    ) +
    theme_bw()
```

## Use ggsector with Seurat

Due to the large raw data of "pbmc", only the code is shown here, but not run. 

Readers are invited to download the data and try it out.

```{r, eval = FALSE}
## Download pbmc data from
# https://cf.10xgenomics.com/samples/cell/pbmc3k/pbmc3k_filtered_gene_bc_matrices.tar.gz
pbmc.data <- Read10X(data.dir = "../filtered_gene_bc_matrices/hg19")
pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)
pbmc <- NormalizeData(pbmc)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
pbmc <- ScaleData(pbmc, features = rownames(pbmc))
pbmc <- RunPCA(pbmc)
pbmc <- RunUMAP(pbmc, dim = 1:10)
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
mks <- tibble::tribble(
    ~type, ~marker,
    "Naive CD4+ T", "IL7R,CCR7",
    "CD14+ Mono", "CD14,LYZ",
    "Memory CD4+", "IL7R,S100A4",
    "B", "MS4A1",
    "CD8+ T", "CD8A",
    "FCGR3A+ Mono", "FCGR3A,MS4A7",
    "NK", "GNLY,NKG7",
    "DC", "FCER1A,CST3",
    "Platelet", "PPBP",
) %>%
    tidyr::separate_rows(marker, sep = ", *") %>%
    dplyr::distinct()
# Dotplot
DotPlot(pbmc, features = unique(mks$marker)) + coord_flip()
# contrast with DotPlot
SectorPlot(pbmc, c(mks$marker, "fsdd"), features_level = unique(rev(mks$marker)))
```

